Synthesizing Symmetric Lenses

### 2018-06-04 papers/SymmetricLenses/main.tex
add my new lens stuff

Critically for this work, \GreedySynth efficiently searches for one of the best
lenses in a class of lenses, characterized by a stochastic regular expression
pair. \GreedySynth works by searching through a smaller search space. First,
\GreedySynth converts the provided stochastic regular expression pair into a
\emph{stochastic DNF regular expression} pair. Where a stochastic regular
expression pair characterizes a given class, a stochastic DNF regular expression
pair is highly normalized -- it is the \emph{unique} pair that characterizes a
given class. Then, we use the syntax of the DNF regular expression pair to
greedily find a highly normalized version of a simple symmetric lens, a
\emph{symmetric DNF lens}. Finally, this symmetric DNF lens is converted into a
simple symmetric lens.

However, these communicating synthesizers pose a problem: unless we find a lens
with no cost, further search may find a better lens. While we may have found a
great lens between two regular expressions, a different regular expression pair
may provide an even better lens. There may be a great lens in the class
characterized by \lstinline{crontab} and \lstinline{scheduled_job}, but it is
not the the intended lens shown in Figure~\ref{fig:example_lens}. However, there
is a tension, we want to find a good lens, but we also want to give quick
feedback to the user. After finding the lens in Figure~\ref{fig:example_lens},
we dont want the algorithm to spend much longer searching for better lenses.

@@

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
    \Leftrightarrow
    \RegexAlt_1 \Concat \RegexAlt_2
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
  \end{tabular}
\end{center}

\afm{there isn't much to say here, should I put these in a figure or something?}

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
    \Leftrightarrow
    \RegexAlt_2 \Concat \RegexAlt_1
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
  \end{tabular}
\end{center}

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_1}$
      \end{cases*}$\\
  \end{tabular}
\end{center}

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \IterateLensOf{\Lens} \OfType
    \StarOf{\Regex}
    \Leftrightarrow
    \StarOf{\RegexAlt}
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String_1\ldots\String_n$
    & =
    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\

    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
    & =
    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\

    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
    & =
    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
      \begin{cases*}
        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
        \Lens.\CreateROf{\String_i} & otherwise
      \end{cases*}$\\
    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
    & =
    & $\String_1'\ldots\String_n'$ where $\String_i' =
      \begin{cases*}
        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
        \Lens.\CreateROf{\StringAlt_i} & otherwise
      \end{cases*}$
  \end{tabular}
\end{center}

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_2}$
      \end{cases*}$\\
  \end{tabular}
\end{center}

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
      \end{cases*}$\\
  \end{tabular}
\end{center}

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Lens_2 \OfType \RegexAlt \Leftrightarrow \RegexAltAlt\\
  }
  {
    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType
    \Regex \Leftrightarrow \RegexAltAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$ & = & $\Lens_2.\CreateROf{(\Lens_1.\CreateROf{\String})}$\\
    $\CreateL{} \App \StringAlt$ & = & $\Lens_1.\CreateLOf{(\Lens_2.\CreateLOf{\StringAlt})}$\\
    $\PutR{} \App \String \App \StringAltAlt$ & = & $\Lens_2.\PutROf{(\Lens_1.\PutROf{\String}{(\Lens_2.\CreateLOf{\StringAltAlt})})}{\StringAltAlt}$\\
    $\PutL{} \App \StringAltAlt \App \String$ & = & $\Lens_1.\PutLOf{(\Lens_2.\PutLOf{\StringAltAlt}{(\Lens_2.\CreateROf{\String})})}{\String}$
  \end{tabular}
\end{center}

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \InvertOf{\Lens} \OfType \RegexAlt \Leftrightarrow \Regex
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
  \end{tabular}
\end{center}
% end ssl

% begin synthesis
\section{Synthesis}
In this work, we aim to quickly find a good lens between two regular expressions
that satisfies the input-output examples. To this end, we develop an information
theoretic metric for how much data is synchronized. Information theoretic
metrics require a notion of probability distributions \emph{how are probability
  distributions being used, are certain things weighted more etc, assign weights
  to different subcomponents and parts of the data}. We annotate
the regular expressions with a reasonable probability distribution, making them
\emph{stochastic regular expressions}. Armed with probability distributions over
the languages, we have the capabilities to define information theoretic metrics
over these probability distributions.

Our overall synthesis algorithm utilizes two synthesizers shown in
Figure~\ref{fig:high-level-algorithm}. There are an infinite number of possible
lenses between two regular expressions that can be searched through. Instead of
searching through all possible lenses, the lenses are grouped into classes of
lenses based on the syntax of the regular expressions they synchronize. Two
lenses are in the same class if they can be typed by the same regular
expressions, using no star equivalences. This works well for synthesis, as
finding a good lens within these classes permits an efficient synthesis
procedure.

However, proposing these classes is complicated by stochastic regular
expressions. While the equivalences on regular expressions are well studied, the
equivalences on stochastic regular expressions are not. We want to propose
equivalent stochastic regular expressions, but both the languages and
probability distributions must be maintained. If the probability distributions
are not maintained, then the metric by which we compare our lenses changes for
each class of lenses, we can't compare lenses between classes. To alleviate
this, we describe the \emph{star-semiring} equivalences on stochastic regular
expressions, and prove they are sound with respect to language and probability
distributions. Our \RXSearch proposes stochastic regular expressions by
traversing a subset of these equivalences, so \RXSearch only proposes regular
expression pairs equivalent to the originals.

The second algorithm, \GreedySynth, searches within a class for the lowest cost
lens.  It is a greedy algorithm, and not guaranteed to find the lowest cost
lens.  However, we haven't found any instances in which \GreedySynth doesn't find
the optimal lens, due to the partial backtracking we implemented. \GreedySynth
works by searching, not through symmetric lenses, but through an alternative
language of Symmetric DNF lenses. After finding such a symmetric DNF lens, that
symmetric DNF lens is converted into a simple symmetric lens.

@@

Much like the synthesis algorithm for Optician, we need to apply rewrites to
handle equivalences on the provided stochastic regular expressions.  However, no
such equivalences have yet been defined.  We extend the \emph{star-semiring}
equivalences to stochastic regular expressions.

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}
% \caption{Regular Expression Equivalences}
% \label{fig:regex-equivalence-rules}

@@

@@

\RXSearch terminates when the number of regular expressions of the smallest
unproposed size exceeds the cost of the current best lens. At this point,
\RXSearch stops proposing regular expression pairs, and instead returns to the
user the best lens found thus far.  Note, if \RXSearch finds a bijective lens (a
lens of cost 0), it will immediately return, as would be expected.

\afm{dont forget to add in description of termination}
%

>>>>

Critically for this work, \GreedySynth efficiently searches for one of the best
lenses in a class of lenses, characterized by a stochastic regular expression
pair. \GreedySynth works by searching through a smaller search space. First,
\GreedySynth converts the provided stochastic regular expression pair into a
\emph{stochastic DNF regular expression} pair. Then, the algorithm uses the
syntax of the DNF regular expression pair to greedily find a highly normalized
version of a simple symmetric lens, a \emph{symmetric DNF lens}. Finally, this
symmetric DNF lens is converted into a simple symmetric lens. DNF Lenses include
neither a composition operator nor a type equivalence rule. Unlike simple
symmetric lenses, for which a single lens term is typed between many possible
regular expression pairs, each DNF lens is only typed between a single regular
expression pair. The relationship between DNF lens syntax and DNF regular
expression syntax is incredibly tight, permitting an incredibly efficient
synthesis algorithm for DNF lenses from DNF regular expressions.

Our strategy involving communicating synthesizers permits both a way of
enumerating classes and efficiently searching through them, but it poses a
problem: unless we find a lens with no cost, other lens classes may contain
better lenses. While we may have found a great lens between two regular
expressions, a different regular expression pair may provide an even better
lens. There may be a great lens in the class characterized by
\lstinline{crontab} and \lstinline{scheduled_job}, but it is not the the
intended lens shown in Figure~\ref{fig:example_lens}. However, there is a
tension, we want to find a good lens, but we also want to give quick feedback to
the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
the algorithm to spend much longer searching for better lenses.

@@

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
    \Leftrightarrow
    \RegexAlt_1 \Concat \RegexAlt_2
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
%    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
%    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
%  \end{tabular}
%\end{center}

Concat is similar to concatenation in existing string lens languages like
Boomerang.  For such terms, we do not provide the semantics, and merely refer
readers to existing work.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
  {
    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
    \Leftrightarrow
    \RegexAlt_2 \Concat \RegexAlt_1
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
%    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
%    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
%  \end{tabular}
%\end{center}
The swap combinator is similar to concat, though the second regular expression
is swapped.

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_1}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \IterateLensOf{\Lens} \OfType
    \StarOf{\Regex}
    \Leftrightarrow
    \StarOf{\RegexAlt}
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
The \IterateLens lens deals with iterated data.

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_2}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \MergeR lens is interesting because it merges data where one data can be in
two formats, and one data has only one format. In previous work, this was
combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
find it more clear to have explicit merge operators: it is easier to see what
lens the synthesis algorithm is creating.

@@

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
      \end{cases*}$\\
  \end{tabular}
\end{center}
The \MergeL lens is symmetric to \MergeR.

\[
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Lens_2 \OfType \RegexAlt \Leftrightarrow \RegexAltAlt\\
  }
  {
    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType
    \Regex \Leftrightarrow \RegexAltAlt
  }
\]
\begin{center}
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$ & = & $\Lens_2.\CreateROf{(\Lens_1.\CreateROf{\String})}$\\
    $\CreateL{} \App \StringAlt$ & = & $\Lens_1.\CreateLOf{(\Lens_2.\CreateLOf{\StringAlt})}$\\
    $\PutR{} \App \String \App \StringAltAlt$ & = & $\Lens_2.\PutROf{(\Lens_1.\PutROf{\String}{(\Lens_2.\CreateLOf{\StringAltAlt})})}{\StringAltAlt}$\\
    $\PutL{} \App \StringAltAlt \App \String$ & = & $\Lens_1.\PutLOf{(\Lens_2.\PutLOf{\StringAltAlt}{(\Lens_2.\CreateROf{\String})})}{\String}$
  \end{tabular}
\end{center}
Composing is interesting in the put functions. Because puts require intermediary
data, we recreate that intermediary data with creates.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
  }
  {
    \InvertOf{\Lens} \OfType \RegexAlt \Leftrightarrow \Regex
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
Inverting is particularly useful when chaining many compositions together, as it
can be used to align the central types.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \equiv \Regex'\\
    \RegexAlt \equiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }
\]
Type equivalence is useful both for addressing type annotations and for
letting compositions go through.
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
Inverting is particularly useful when chaining many compositions together, as it
can be used to align the central types.
% end ssl

% begin synthesis
\section{Synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm. The
provided regular expressions are first converted into stochastic regular
expressions with \ToStochastic. Next, $\RXSearch$ is initialized with that pair
of stochastic regular expressions. $\RXSearch$ then proposes candidate classes
of lenses by proposing regular expression pairs that characterize those classes.
$\GreedySynth$ greedily finds a low cost lens within each proposed lens class,
if there is one that satisfies the specification. These lenses are then provided
to $\RXSearch$. Eventually, $\RXSearch$ is content with one of the lenses, and
returns that lens to the user.

\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\Regex,\RegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\Regex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Init}{\Regex,\RegexAlt}$
    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Consider}{\RXSearchState,\LCO}$
    \EndWhile
    \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

%In this work, we aim to quickly find a good lens between two regular expressions
%that satisfies the input-output examples. To this end, we develop an information
%theoretic metric for how much data is synchronized. Information theoretic
%metrics require a notion of probability distributions \emph{how are probability
%  distributions being used, are certain things weighted more etc, assign weights
%  to different subcomponents and parts of the data}. We annotate the regular
%expressions with a reasonable probability distribution, making them
%\emph{stochastic regular expressions}. Armed with probability distributions over
%the languages, we have the capabilities to define information theoretic metrics
%over these probability distributions.
%
%Our overall synthesis algorithm utilizes two synthesizers shown in
%Figure~\ref{fig:high-level-algorithm}. There are an infinite number of possible
%lenses between two regular expressions that can be searched through. Instead of
%searching through all possible lenses, the lenses are grouped into classes of
%lenses based on the syntax of the regular expressions they synchronize. Two
%lenses are in the same class if they can be typed by the same regular
%expressions, using no star equivalences. This works well for synthesis, as
%finding a good lens within these classes permits an efficient synthesis
%procedure.
%
%However, proposing these classes is complicated by stochastic regular
%expressions. While the equivalences on regular expressions are well studied, the
%equivalences on stochastic regular expressions are not. We want to propose
%equivalent stochastic regular expressions, but both the languages and
%probability distributions must be maintained. If the probability distributions
%are not maintained, then the metric by which we compare our lenses changes for
%each class of lenses, we can't compare lenses between classes. To alleviate
%this, we describe the \emph{star-semiring} equivalences on stochastic regular
%expressions, and prove they are sound with respect to language and probability
%distributions. Our \RXSearch proposes stochastic regular expressions by
%traversing a subset of these equivalences, so \RXSearch only proposes regular
%expression pairs equivalent to the originals.
%
%The second algorithm, \GreedySynth, searches within a class for the lowest cost
%lens.  It is a greedy algorithm, and not guaranteed to find the lowest cost
%lens.  However, we haven't found any instances in which \GreedySynth doesn't find
%the optimal lens, due to the partial backtracking we implemented. \GreedySynth
%works by searching, not through symmetric lenses, but through an alternative
%language of Symmetric DNF lenses. After finding such a symmetric DNF lens, that
%symmetric DNF lens is converted into a simple symmetric lens.

@@

$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
However, as of yet, there is no syntactic means to reason about stochastic
regular expression equivalence. We extend the \emph{star-semiring} equivalences
to stochastic regular expressions.

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

@@

Finally, we wish to infer a reasonable probability distribution for stochastic
regular expressions.  While these probability distributions could be either
input by users or inferred from example data, we found that our inferred
probability distributions work in practice. We defer our description of
\ToStochastic until \S\ref{sec:dnf}, as it is easier to describe our desired
probability distribution in the context of DNF regular expressions.
% \caption{Regular Expression Equivalences}
% \label{fig:regex-equivalence-rules}

@@

\RXSearch terminates when the number of regular expressions of the smallest
unproposed size exceeds the cost of the current best lens. At this point,
\RXSearch stops proposing regular expression pairs, and instead returns to the
user the best lens found thus far. Note this means that if \RXSearch finds a
bijective lens (a lens of cost 0) it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions are used as a highly normalized form of
stochastic regular expressions. Due to this highly normalized nature, where type
equivalence is often needed for simple symmetric lenses, such rules aren't
needed when typed between DNF regular expressions, as many semantically equal
but syntactically distinct regular expressions have syntactically equal DNF
forms.
